Sperm should evolve to make female meiosis fair

### 2014-05-21 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
yb changed replaced refs to Table S1 with refs to Dataset S1 because datasets and not tables can be Excel format

\section{Results}
\subsection{Invasion of an allele that promotes its own drive}
In the standard single-locus, biallelic model of female meiotic drive,
the driving allele is transmitted to the egg in heterozygotes with
probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004} and see Model 1 in the Appendix for more details). 
To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Table S1). 
We then identify the conditions allowing for the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
We conclude our single locus results with an analysis of a related model (Model 3) - in which
drivers influence their transmission in females via paternal genotype,
rather than sperm haplotype.

@@

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Table S1). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Appendix S1).

@@

\section{Appendix}
We highlight simple analytical results from: 
	the standard model of female meiotic drive (Model 1); a sperm-dependent 
	model of female meiotic drive (Model 2); a male-genotype-dependent model of
	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
	respectively, while more complex models include more loci and/or alleles.  
Transmission rules describing the outcomes of all matings in each
	model are presented in Table S1. 
The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. 
Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
	and $f_g''$ in adults after selection. 
After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
We verify these approximations with exact numerical iterations in Figures S1-S3 , and detail the derivation of these approximations in Appendix S1.

>>>>

\section{Results}
\subsection{Invasion of an allele that promotes its own drive}
In the standard single-locus, biallelic model of female meiotic drive,
the driving allele is transmitted to the egg in heterozygotes with
probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004} and see Model 1 in the Appendix for more details). 
To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Dataset S1). 
We then identify the conditions allowing for the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
We conclude our single locus results with an analysis of a related model (Model 3) - in which
drivers influence their transmission in females via paternal genotype,
rather than sperm haplotype.

@@

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Dataset S1). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Appendix S1).

@@

\section{Appendix}
We highlight simple analytical results from: 
	the standard model of female meiotic drive (Model 1); a sperm-dependent 
	model of female meiotic drive (Model 2); a male-genotype-dependent model of
	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
	respectively, while more complex models include more loci and/or alleles.  
Transmission rules describing the outcomes of all matings in each
	model are presented in Dataset S1. 
The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. 
Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
	and $f_g''$ in adults after selection. 
After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
We verify these approximations with exact numerical iterations in Figures S1-S3 , and detail the derivation of these approximations in Appendix S1.

### 2014-05-21 writeup_and_notes/pnasSubmission/Text_S1.tex
yb changed replaced refs to Table S1 with refs to Dataset S1 because datasets and not tables can be Excel format

\section*{SI Table legends}
  \subsection*{Table S1}
  We detail the expected frequencies of each zygote from a given mating in Table S1. 
  Table S1 is an Excel Spreadsheet with six tabs - each one corresponding to Models 1-6. 
  Table S1 can be used to generate genotype frequencies after drive and selection, 
  by summing the  frequency of zygotes produced by a given mating weighted by the frequency of this mating. 
  Assuming random mating this weighting is simply the product of maternal and paternal genotype frequencies after selection in the previous generation.

>>>>

\section*{SI Table legends}
  \subsection*{Dataset S1}
  We detail the expected frequencies of each zygote from a given mating in Dataset S1. 
 Dataset S1 is an Excel Spreadsheet with six tabs - each one corresponding to Models 1-6. 
 Dataset S1 can be used to generate genotype frequencies after drive and selection, 
  by summing the  frequency of zygotes produced by a given mating weighted by the frequency of this mating. 
  Assuming random mating this weighting is simply the product of maternal and paternal genotype frequencies after selection in the previous generation.
